---
author: "Anna Iashina"
title: "Behavior and network imputation"
output: html_document
---

```{r}
siena07ToConvergence <- function(alg, dat, eff, ans0=NULL, threshold, ...){
  numr <- 0
  ans <- siena07(alg, data = dat, effects = eff, prevAns = ans0, nbrNodes = 4, returnDeps = TRUE, useCluster = TRUE, ...) # the first run
  repeat {
    save(ans, file = paste("ans",numr,".RData",sep = "")) # to be safe
    numr <- numr + 1           # count number of repeated runs
    tm <- ans$tconv.max      # convergence indicator
    cat(numr, tm,"\n")       # report how far we are
    if (tm < threshold) {break}   # success
    if (tm > 10) {stop()}     # divergence without much hope
    # of returning to good parameter values
    if (numr > 100) {stop()}  # now it has lasted too long
    alg$nsub <- 1
    alg$n2start <- 5000 + numr * 1000
    ans <- siena07(alg, data = dat, effects = eff, prevAns = ans,nbrNodes = 4, returnDeps = TRUE, useCluster = TRUE, ...)
  }
  if (tm > threshold)
  {
    stop("Warning: convergence inadequate.\n")
  }
  ans
}

library(RSienaTest)
library(mice)

```


### Data preprocessing

Glasgow data contains natural missing data. For the research purposes we need a complete dataset so we leave only those observations which doesn't contain any missings.

```{r}

# Read data sets:
load("~/Desktop/siena_project/data/Glasgow-demographic.RData")# for gender
load("~/Desktop/siena_project/data/Glasgow-friendship.RData") # for networks
load("~/Desktop/siena_project/data/Glasgow-substances.RData") # for alcochol and tobacco
load("~/Desktop/siena_project/data/Glasgow-selections.RData") # for 129 selection
selection129 = selection129

# firstly we leave only 129 students who don't have missings in tobacco variable
alcohol <- alcohol[selection129,]
tobacco <- tobacco[selection129,]

sex.F <- sex.F[selection129]

friendship.1 <- friendship.1[selection129, selection129]
friendship.2 <- friendship.2[selection129, selection129]

# change coding
friendship.1[friendship.1 == 2] <- 1
friendship.2[friendship.2 == 2] <- 1

# 129 selection has missings in alcochol variable. Let's find those who also has no missings in alcochol variable
present <- !is.na(rowSums(alcohol))

alcohol <- alcohol[present,]
tobacco <- tobacco[present,]

sex.F <- sex.F[present]

friendship.1 <- friendship.1[present, present]
friendship.2 <- friendship.2[present, present]

N <- nrow(friendship.1) # number of observations

```


### SIENA model with full data

Before we proceed to the imputation models. we need to make sure the model we are going to estimate is correct. We'll estimate the model with behavioral effects with the full data we have. Then we'll take a look on goodness of fit.


```{r}

friendship <- sienaDependent(array(c(friendship.1, friendship.2),
                                     dim = c( N, N, 2)))# create dependent variable

# for behavior effects 
drinkingbeh <- sienaDependent(alcohol[,1:2], type = "behavior" ) # complete data
smokingbeh <- sienaDependent(tobacco[,1:2], type = "behavior" )

# adding gender covariate
sex <- coCovar(sex.F)

# create siena data
myCoEvolutionData <- sienaDataCreate(friendship, drinkingbeh,
                                     smokingbeh,
                                     sex)

myCoEvolutionEff <- getEffects(myCoEvolutionData) # get initial effects

myCoEvolutionEff <- includeEffects(myCoEvolutionEff, density, recip,  inPopSqrt, cycle3, transTrip, transRecTrip, inActSqrt) #structural effects

#myCoEvolutionEff <- setEffect(myCoEvolutionEff, outTrunc, parameter = 7, test = FALSE, fix = TRUE,              initialValue = -10) # we cannot use MaxDegree when running MoM estimation, so we use outTrunc instead

# gender homophily
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, sameX, interaction1 = "sex" )

# selection effects
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, egoX, altX, egoXaltX, interaction1 = "drinkingbeh" )
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, egoX, altX, egoXaltX, interaction1 = "smokingbeh" )

myCoEvolutionEff <- includeEffects(myCoEvolutionEff, name = "drinkingbeh", avAlt, interaction1 = "friendship")
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, name = "smokingbeh", avAlt, interaction1 = "friendship")

# gender influence on smoking and drinking (not for the first wave)
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                         name = "smokingbeh", interaction1 = "sex")
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                       name = "drinkingbeh", interaction1 = "sex")

# smoking and drinking influencing each other (not for the first wave)
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                   name = "smokingbeh", interaction1 = "drinkingbeh")
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                   name = "drinkingbeh", interaction1 = "smokingbeh")

# check the list of effects
myCoEvolutionEff


# Estimate the siena model
estimation.options <- sienaAlgorithmCreate(seed = 846, n3 = 3000) # seed must be specified

# estimate the model until it converges
model.1 <- siena07ToConvergence2(alg = estimation.options,
                                dat = myCoEvolutionData,
                                eff = myCoEvolutionEff,
                                threshold = .25, nodes = 2)


# make a tex table
siena.table(model.1, type = "tex", sig = TRUE)

model.1
```

Now, when we obtained the full model, we ought to test the goodness of fit.

### Goodness of fit

Note: *networks* and *sna* packages must be installed

We'll check the goodness of indegree and outdegree distribution and the number of different triads. The violin plots let us compare empirical distribution (red line) with simulated distributions of those parameters and triads (violins). 

```{r}
# Indegree
gof1.id <- sienaGOF(model.1, verbose = TRUE,
                    varName = "friendship", IndegreeDistribution, cumulative = FALSE)
plot(gof1.id)


# Outdegree
gof1.od <- sienaGOF(model.1, verbose = TRUE, varName = "friendship",
                    OutdegreeDistribution, cumulative = FALSE)
plot(gof1.od)


# Triad census

TriadCensus <- function(i, data, sims, wave, groupName, varName, levls=1:16){
  unloadNamespace("igraph") # to avoid package clashes
  require(sna)
  require(network)
  x <- networkExtraction(i, data, sims, wave, groupName, varName)
  if (network.edgecount(x) <= 0) {x <- symmetrize(x)}
  # because else triad.census(x) will lead to an error
  tc <- sna::triad.census(x)[1,levls]
  # triad names are transferred automatically
  tc
}


gof1.tc <- sienaGOF(model.1, verbose = TRUE,
                    varName = "friendship", TriadCensus)
# This is more time-consuming...

plot(gof1.tc, scale = TRUE, center = TRUE)

```



#######################################################
# Multiple imputation for behavioral and network data #
#######################################################


### Generating missing data

We will generate 20% of missings (MCAR) for the behavioral data (alcochol and tobacco consumption) and for the networks. The missing cases are persistent by waves. 

```{r}

set.seed(2509609) # for consistency

# generate the random samples for each wave
missW1 <- sample(1:N,21)
missW2 <- sample(1:N,21)

# make copies of the original data
f1miss <- friendship.1
f2miss <- friendship.2

alcMiss <- alcohol
tobMiss <- tobacco

# create missings
f1miss[missW1,] <- NA # networks
f2miss[missW2,] <- NA

alcMiss[missW1, 1] <- NA # alcochol
alcMiss[missW2, 2] <- NA

tobMiss[missW1, 1] <- NA # tobacco
tobMiss[missW2, 2] <- NA

```

### MICE imputation for the behavior variables

For the alcochol and tobacco imputation in the 1st wave we use indegree in the 1st and 2nd waves, average alter efect on alcochol and tobacco. Alcochol and tobacco variable are also used to help impute each other.

```{r}
indegree1 <- colSums(f1miss, na.rm = TRUE) # count indegree 1st wave
indegree2 <- colSums(f2miss, na.rm = TRUE) # count indegree 2nd wave

# split the behavior variables by waves
a1 = alcMiss[,1]
t1 = tobMiss[,1]
a2 = alcMiss[,2]
t2 = tobMiss[,2]

# calculate the avgAlt effect where applicable (add a formula here)

avgAltA1 <- rowSums(sweep(t(friendship.1), MARGIN = 2, a1,'*'),na.rm = TRUE) / rowSums(t(friendship.1), na.rm = TRUE)
avgAltT1 <- rowSums(sweep(t(friendship.1), MARGIN = 2, t1,'*'),na.rm = TRUE) / rowSums(t(friendship.1), na.rm = TRUE)

avgAltA1[is.nan(avgAltA1)] <- NA
avgAltT1[is.nan(avgAltT1)] <- NA

avgAltA2 <- rowSums(sweep(t(friendship.2), MARGIN = 2, a2,'*'),na.rm = TRUE) / rowSums(t(friendship.2), na.rm = TRUE)
avgAltT2 <- rowSums(sweep(t(friendship.2), MARGIN = 2, t2,'*'),na.rm = TRUE) / rowSums(t(friendship.2), na.rm = TRUE)

avgAltA2[is.nan(avgAltA2)] <- NA
avgAltT2[is.nan(avgAltT2)] <- NA

# create the final dataframe for mice imputation
df <- cbind(a1, t1, sex.F,a2, t2, indegree1, indegree2, avgAltA1, avgAltT1, avgAltA2, avgAltT2)

```

With mice imputation all the variables that contain missings will be imputed but we are only interested in the first wave alcochol and tobacco imputation. The imputation for the behavioral variables in the second wave will be done with SAOMs later. 

```{r message=FALSE, warning=FALSE, include=FALSE}

D <- 50 # MUST BE SET TO A HIGHER VALUE
# impute behavior wave 1
m = mice(df, m = D ,defaultMethod = "pmm", maxit = 100)

```

### First imputation with stationary SAOM


Now, when we have the imputed data for the behavioral variables in the 1st wave, we can start imputing the network. The first run of imputation for the 1st wave network is done with stationary SAOM. We take one of the behavioral datasets completed with mice and use it to model the first wave sructure.

Since version 1.2-12, Maximum Likelihood (ML) estimation by $\textsf{Rsiena}$ with ```returnDeps = TRUE``` returns an edgelist of the final network at the end of the phase 3 simulation. The following function, ``` getNet() ```, uses this edgelist to impute the data.

```{r}

getNet <- function(observedNet,edgeList) { # get imputed network from SIENA model
  # observedNet = observed network as adjacency with missing data
  # edgeList = edgelist that is returned by siena07(...)$sims
  observedNet[is.na(observedNet)] <- 0
  for (i in 1:nrow(edgeList)) {
    observedNet[edgeList[i,1],edgeList[i,2]] <- 1
  }
  return(observedNet)
}


```



```{r}

# the model

friendship <- sienaDependent(array(c(f1miss, f1miss), dim = c( N, N, 2)), # make a dependent variable
                            allowOnly = FALSE )
  
completedData <- complete(m, sample(1:D, 1)) # a random behavior dataset imputed with MICE

drinking <- coCovar(completedData$a1) # imputed alco w1
smoking <- coCovar(completedData$t1) # imputed tobacco w1
sex <- coCovar(sex.F)
  
w2 <- coDyadCovar(f2miss) # the 2nd wave incomplete network as covariate 
  
Data.stationary  <- sienaDataCreate(friendship, drinking, smoking, sex, w2)
  
effects.stationary <- getEffects(Data.stationary)

effects.stationary <- includeEffects(effects.stationary, density, recip, inPopSqrt,
                                     cycle3, transTrip, transRecTrip, inActSqrt)
                                
# gender homophily
effects.stationary <- includeEffects(effects.stationary, sameX, interaction1 = "sex" )


effects.stationary <- includeEffects(effects.stationary, egoX, altX, egoXaltX,
                                     interaction1 = "drinking")
effects.stationary <- includeEffects(effects.stationary, egoX, altX, egoXaltX,
                                     interaction1 = "smoking")

# 2nd wave as coDyad covariate    
effects.stationary <- includeEffects(effects.stationary, X, name = "friendship",
                                     interaction1 = "w2")

# fix the rate
effects.stationary <- setEffect(effects.stationary, Rate, initialValue = 20,
                                name = "friendship", fix = TRUE, type = "rate") 

estimation.options <- sienaAlgorithmCreate(useStdInits = FALSE,
                                           seed = 214,
                                           n3 = 1000, maxlike = FALSE,
                                           cond = FALSE#, MaxDegree = c(friendship = 6)
                                           )

# First estimate the stationary SAOM by MoM.
# ~ 60 mins
period0saom <- siena07ToConvergence(alg = estimation.options,
                                    dat = Data.stationary,
                                    eff = effects.stationary, threshold = 0.25, nodes = 2)

```


### Multiple SAOM imputation starting from the first estimate by MoM with the stationary SAOM

```{r message=FALSE, warning=FALSE}

set.seed(142)

D <- D # the same number as for the mice imputation
wave1imp <- list()


for (j in 1:D) {
  cat('imputation', j,'\n')
  # We have to change one tie in the starting network from 1 to 0 to allow ML
  # simulation. For every imputation we want to change a different tie.
  # Further, we will fix all observed ties and non-ties, which means we convert
  # them to structural zeros (10) and structural ones (11). This increases speed
  # and reduces randomness in our imputation. However, this is only done in the
  # imputation of the first wave. We do not want observed ties to change during
  # the stationary SAOM simulation. In contrast, in later waves observed ties
  # are likely to change and this change needs to be preserved.
  
  n1 <- f1miss
  n1 <- n1 + 10
  diag(n1) <- 0
  n2 <- n1
  tieList <- c(1:(nrow(n1)**2))[c(n1 == 11)]
  tieList <- tieList[!is.na(tieList)]
  
  changedTie <- sample(tieList,1)
  
  n1[changedTie] <- 0
  n2[changedTie] <- 1
  
  friendship <- sienaDependent(array(c(n1,n2), dim = c( N, N, 2)),
                            allowOnly = FALSE )
  
  
  completedData <- complete(m, j) # j-completed dataset from MICE
  
  drinking <- coCovar(completedData$a1) # alco w1
  smoking <- coCovar(completedData$t1) # tobacco w1
  sex <- coCovar(sex.F)
  
  w2 <- coDyadCovar(f2miss) # 2nd wave as coDyad covariate
  
  Data.stationary  <- sienaDataCreate(friendship, drinking, smoking, sex, w2)
  
  effects.stationary <- getEffects(Data.stationary)
  effects.stationary <- includeEffects(effects.stationary, density, recip,
                                         #outTrunc,
                                         inPopSqrt, cycle3, transTrip, transRecTrip, inActSqrt)
  
  #effects.stationary <- setEffect(effects.stationary, outTrunc, parameter = 7, test = FALSE, fix = TRUE, initialValue = -10) # we        cannot use MaxDegree when running MoM estimation, so we use outTrunc instead
  
  # gender homophily
  effects.stationary <- includeEffects(effects.stationary, sameX, interaction1 = "sex" )
  

  effects.stationary <- includeEffects(effects.stationary, egoX, altX, egoXaltX,
                                     interaction1 = "drinking")
  effects.stationary <- includeEffects(effects.stationary, egoX, altX, egoXaltX,
                                     interaction1 = "smoking")
    
  effects.stationary <- includeEffects(effects.stationary, X, name = "friendship",
                                     interaction1 = "w2")

  # set fixed rate
  effects.stationary <- setEffect(effects.stationary, Rate, initialValue = 20, 
                                name = "friendship", fix = TRUE, type = "rate")
  
  imputation.options <- sienaAlgorithmCreate(useStdInits = FALSE, seed = 13848, # always specify seed
                                           n3 = 10, maxlike = TRUE,
                                           cond = FALSE, nsub = 0,
                                           simOnly = TRUE) # no MaxDegree here because we use outTrunc instead
  
  # run the estimation
  sims <- siena07(imputation.options, data = Data.stationary,
                  effects = effects.stationary,
                  prevAns = period0saom,
                  returnDeps = TRUE)$sims[[10]]

  
  wave1imp[[j]] = getNet(f1miss, sims[[1]]) # save every imputed network
  
}


```



### 2nd wave imputation

Imputation of missings in wave 2 by SAOM MoM simulations

After obtaining D imputations of wave 1 we proceed by imputing wave 2 in a stepwise procedure.


```{r W2 imputation, message=FALSE, include=FALSE}
wave2imp <- list()
alcImpW2 <- matrix(NA,nrow(friendship.1),D)
tobImpW2 <- matrix(NA,nrow(friendship.1),D)

for (j in 1:D) {
  
  cat('imputation',j,'\n')

  # now impute wave2
  friendship <- sienaDependent(array(c(wave1imp[[j]], f2miss),
                                  dim = c(N,N,2)))
  
  completedData <- complete(m, j)
  
  alc <- cbind(completedData$a1, completedData$a2) # what if we try a2 with missings here
  tob <- cbind(completedData$t1, completedData$t2) # what if we try t2 with missings here
  
  drinkingbeh <- sienaDependent(alc, type = "behavior") 
  smokingbeh <- sienaDependent(tob, type = "behavior")
  
  sex = coCovar(sex.F)
  
  Data.w2  <- sienaDataCreate(friendship, drinkingbeh, smokingbeh, sex)
  
  effects.twoWaves <- getEffects(Data.w2)
  
  effects.twoWaves <- includeEffects(effects.twoWaves, density, recip, inPopSqrt,
                                     cycle3, transTrip, transRecTrip, inActSqrt)
  
  # gender homophily
  effects.twoWaves <- includeEffects(effects.twoWaves, sameX, interaction1 = "sex" )
  
  effects.twoWaves <- includeEffects(effects.twoWaves, egoX, altX, egoXaltX,  
                                     interaction1 =  "drinkingbeh")
  effects.twoWaves <- includeEffects(effects.twoWaves, egoX, altX, egoXaltX,
                                     interaction1 =  "smokingbeh")
  
  effects.twoWaves <- includeEffects(effects.twoWaves, name = "drinkingbeh", avAlt, interaction1 = "friendship")
  effects.twoWaves <- includeEffects(effects.twoWaves, name = "smokingbeh", avAlt, interaction1 = "friendship")

  # gender influence on smoking and drinking 
  effects.twoWaves <- includeEffects(effects.twoWaves, effFrom,
                                         name = "smokingbeh", interaction1 = "sex")
  effects.twoWaves <- includeEffects(effects.twoWaves, effFrom,
                                       name = "drinkingbeh", interaction1 = "sex")

  # smoking and drinking influencing each other 
  effects.twoWaves <- includeEffects(effects.twoWaves, effFrom,
                                   name = "smokingbeh", interaction1 = "drinkingbeh")
  effects.twoWaves <- includeEffects(effects.twoWaves, effFrom,
                                   name = "drinkingbeh", interaction1 = "smokingbeh")
  
  
  estimation.options <- sienaAlgorithmCreate(useStdInits = FALSE,
                                           seed = 2214,
                                           n3 = 1000, maxlike = FALSE,
                                           cond = FALSE#, MaxDegree = c(friendship = 6)
                                           )
  

  if (j == 1) {
    period1saom <- siena07ToConvergence2(alg = estimation.options,
                                        dat = Data.w2,
                                        eff = effects.twoWaves,
                                        threshold = 0.25, nodes = 2)
  } else {
    period1saom <- siena07ToConvergence2(alg = estimation.options,
                                        dat = Data.w2,
                                        eff = effects.twoWaves,
                                        threshold = 0.25,
                                        ans0 = period1saom, nodes = 2)
  }
  
  # we want to impute the 2nd wave variables with SAOMs
  alc <- cbind(completedData$a1, a2)
  tob <- cbind(completedData$t1, t2)
  
  drinkingbeh <- sienaDependent(alc, type = "behavior")
  smokingbeh <- sienaDependent(tob, type = "behavior")
  
  Data.w2  <- sienaDataCreate(friendship, drinkingbeh, smokingbeh, sex)
  
  sims <- siena07(imputation.options, data = Data.w2,
                  effects = effects.twoWaves,
                  prevAns = period1saom,
                  returnDeps = TRUE)$sims[[10]]
  
  wave2imp[[j]] <- getNet(f2miss, sims[[1]]) # save imputed 2nd wave networks

  alcImpW2[,j] = sims[[2]] # save imputed alco
  tobImpW2[,j] = sims[[3]] # save imputed tobacco
  
}


save.image('mi.RData') 
# Now that we have obtained our imputed data sets we need to run our model
# on the D complete data sets.

```

############################
# Imputed model estimation #
############################


```{r imputed model estimation, message=FALSE, include=FALSE}
saomResults <- list()

for (j in D:1) {
  cat('Imputation',j,'\n')
  
  friendship <- sienaDependent(array(c(wave1imp[[j]], wave2imp[[j]]), dim = c( N, N, 2)))
  
  #if (j == 1) {
   # completedData <- complete(m, j)
   # alc = cbind(completedData$a1, completedData$a2)
  #  tob = cbind(completedData$t1, completedData$t2)
 # } else {
    completedData <- complete(m, j)
    alc = cbind(completedData$a1, alcImpW2[,j])
    tob = cbind(completedData$t1, tobImpW2[,j])
 # }
  
  
  drinkingbeh <- sienaDependent(alc, type = "behavior")
  smokingbeh <- sienaDependent(tob, type = "behavior")
  
  sex <- coCovar(sex.F)
  
  Data  <- sienaDataCreate(friendship, drinkingbeh, smokingbeh, sex)
  
  effectsData <- getEffects(Data)
  effectsData <- includeEffects(effectsData, density, recip, inPopSqrt, cycle3,
                                transTrip, transRecTrip, inActSqrt)
  
  # gender homophily
  effectsData <- includeEffects(effectsData, sameX, interaction1 = "sex" )
  
  effectsData <- includeEffects(effectsData, egoX, altX, egoXaltX,  
                                     interaction1 =  "drinkingbeh")
  effectsData <- includeEffects(effectsData, egoX, altX, egoXaltX,
                                     interaction1 =  "smokingbeh")
  
  effectsData <- includeEffects(effectsData, name = "drinkingbeh", avAlt, interaction1 = "friendship")
  effectsData <- includeEffects(effectsData, name = "smokingbeh", avAlt, interaction1 = "friendship")

  # gender influence on smoking and drinking 
  effectsData <- includeEffects(effectsData, effFrom,
                                         name = "smokingbeh", interaction1 = "sex")
  effectsData <- includeEffects(effectsData, effFrom,
                                       name = "drinkingbeh", interaction1 = "sex")

  # smoking and drinking influencing each other 
  effectsData <- includeEffects(effectsData, effFrom,
                                   name = "smokingbeh", interaction1 = "drinkingbeh")
  effectsData <- includeEffects(effectsData, effFrom,
                                   name = "drinkingbeh", interaction1 = "smokingbeh")
  
  estimation.options <- sienaAlgorithmCreate(useStdInits = FALSE, seed = 38712,
                                             n3 = 5000, maxlike = FALSE,
                                             lessMem = TRUE#, MaxDegree = c(friendship=max(rowSums(wave1imp[[j]])))
                                             )
  if (j == 50) {
    saomResults[[j]] <- siena07ToConvergence(alg = estimation.options,
                                             dat = Data, eff = effectsData,
                                             threshold = 0.25, nodes = 4)
  } else {
    saomResults[[j]] <- siena07ToConvergence(alg = estimation.options,
                                             dat = Data, eff = effectsData,
                                             threshold = 0.25,
                                             ans0 = saomResults[[j + 1]], nodes = 4)
  }
}


```


###### Default SIENA imputation model (not working now)

```{r}
friendship <- sienaDependent(array(c(f1miss, f2miss),
                                     dim = c( N, N, 2)))# create dependent variable

# for behavior effects 
drinkingbeh <- sienaDependent(alcMiss[,1:2], type = "behavior" ) # complete data
smokingbeh <- sienaDependent(tobMiss[,1:2], type = "behavior" )

# adding gender covariate
sex <- coCovar(sex.F)

# create siena data
myCoEvolutionData <- sienaDataCreate(friendship, drinkingbeh,
                                     smokingbeh,
                                     sex)

myCoEvolutionEff <- getEffects(myCoEvolutionData) # get initial effects

myCoEvolutionEff <- includeEffects(myCoEvolutionEff, density, recip,  inPopSqrt, cycle3, transTrip, transRecTrip, inActSqrt) #structural effects

# gender homophily
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, sameX, interaction1 = "sex" )

# selection effects
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, egoX, altX, egoXaltX, interaction1 = "drinkingbeh" )
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, egoX, altX, egoXaltX, interaction1 = "smokingbeh" )

myCoEvolutionEff <- includeEffects(myCoEvolutionEff, name = "drinkingbeh", avAlt, interaction1 = "friendship")
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, name = "smokingbeh", avAlt, interaction1 = "friendship")

# gender influence on smoking and drinking (not for the first wave)
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                         name = "smokingbeh", interaction1 = "sex")
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                       name = "drinkingbeh", interaction1 = "sex")

# smoking and drinking influencing each other (not for the first wave)
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                   name = "smokingbeh", interaction1 = "drinkingbeh")
myCoEvolutionEff <- includeEffects(myCoEvolutionEff, effFrom,
                                   name = "drinkingbeh", interaction1 = "smokingbeh")

# check the list of effects
myCoEvolutionEff


# Estimate the siena model
estimation.options <- sienaAlgorithmCreate(seed = 8146, n3 = 3000) # seed must be specified

# estimate the model until it converges
model.2 <- siena07ToConvergence(alg = estimation.options,
                                dat = myCoEvolutionData,
                                eff = myCoEvolutionEff,
                                threshold = .25)


# make a tex table
siena.table(model.2, type = "tex", sig = TRUE)

model.2
```




```{r}
# Now we have D RSiena results and all that is left is to combine them

# We need to extract all parameter and standard error estimates from the models
# and combine them using Rubin's Rules.

npar <- sum(effectsData$include)

rowVar <- function(x) {
  rowSums((x - rowMeans(x))^2)/(dim(x)[2] - 1)
}

MIResults <- as.data.frame(matrix(NA,npar,(2 * D)))

for (i in 1:D) {
  names(MIResults)[i * 2 - 1] <- paste("imp" , "mean", sep = as.character(i))
  names(MIResults)[i * 2] <- paste("imp" , "se", sep = as.character(i))
  MIResults[,i * 2 - 1] <- saomResults[[i]]$theta  # estimates
  MIResults[,i * 2] <-  sqrt(diag(saomResults[[i]]$covtheta))  # standard errors
}

# Now we get the average covariance structure between the parameters
WDMIs <- matrix(0,npar,npar)

for (i in 1:D) {
  WDMIs <- WDMIs + saomResults[[i]]$covtheta
}

WDMIs <- (1/D) * WDMIs

# Using Rubin's Rules we combine the parameters and standard errors and
# complete the procedure
finalResults <- as.data.frame(matrix(NA,npar,2))
names(finalResults) <- c("combinedEstimate", "combinedSE")
rownames(finalResults) <- effectsData$effectName[effectsData$include]

finalResults$combinedEstimate <- rowMeans(MIResults[,seq(1,2*D,2)])
finalResults$combinedSE <- sqrt(diag(WDMIs) + ((D + 1)/D) *
                                  rowVar(MIResults[,seq(1,2*D,2)]))

finalResults$completeTheta <- model.1$theta
finalResults$completeSE <- sqrt(diag(model.1$covtheta))

finalResults$defaultTheta <- model.2$theta
finalResults$defaultSE <- sqrt(diag(model.2$covtheta))


round(finalResults, 4)
# the names of user specified interactions have to be set manually

print(xtable(round(finalResults, 4), type = "latex"), file = "BdTdRound.tex")

save.image('mi.RData') # 100 MB
```



```{r}
###############################################################################
# proportion of variance due to imputation
BdTd <- as.data.frame(matrix(,npar,1))

names(BdTd) <- c("parameter")

BdTd$parameter <- finalResults$parameter

BdTd$MIsaom <- round(((51/50) * rowVar(MIResults[,seq(1,100,2)])) / (finalResults$combinedSE ** 2),2)

View(BdTd)

print(xtable(BdTd, type = "latex"), file = "BdTdRound.tex")

```


```{r}
################################################################################
#
# Plots
#
finalResults$parameter <- c("Friend rate 1",
                          "Density",
                          "Reciprocity",
                          "transTrip",
                          "transRecTrip",
                          "3-cycles",
                          "Indeg. Pop. sqrt.",
                          "Indeg. Act. sqrt.",
                          "Same sex",
                          "Alter drinking",
                          "Ego drinking",
                          "Ego x alt drinking",
                          "Alter smokimg",
                          "Ego smoking",
                          "Ego x alt smoking",
                          "Drinking rate 1",
                          "Drink. linear",
                          "Drink. quadratic",
                          "Avg. Alter Drink.",
                          "Drink: eff. from sex",
                          "Drink: eff. from smoke.",
                          "Smoking rate 1",
                          "Smoke. linear",
                          "Smoke. quadratic",
                          "Avg. Alter Smoke.",
                          "Smoke: eff. from sex",
                          "Smoke: eff. from drink.")



# lognformat - dirty

ld <- as.data.frame(matrix(NA,81,3))

ld[,1] <- rep(finalResults$parameter,3)
ld[1:27,c(2,3)] <- finalResults[,c(1,2)]
ld[28:54,c(2,3)] <- finalResults[,c(3,4)]
ld[55:81,c(2,3)] <- finalResults[,c(5,6)]


ld[,4] <- c(rep("MI-SAOM",27),
            rep("Complete Data",27),
            rep("Default SIENA",27)
            )

names(ld) <- c("parameter","Theta","SE","model")

ld$model = factor(ld$model, levels = c("MI-SAOM", "Default SIENA", "Complete Data" )[3:1]) # ,"Default SIENA"

ld$parameter = factor(ld$parameter,
                          levels = c("Friend rate 1",
                          "Density",
                          "Reciprocity",
                          "transTrip",
                          "transRecTrip",
                          "3-cycles",
                          "Indeg. Pop. sqrt.",
                          "Indeg. Act. sqrt.",
                          "Same sex",
                          "Alter drinking",
                          "Ego drinking",
                          "Ego x alt drinking",
                          "Alter smokimg",
                          "Ego smoking",
                          "Ego x alt smoking",
                          "Drinking rate 1",
                          "Drink. linear",
                          "Drink. quadratic",
                          "Avg. Alter Drink.",
                          "Drink: eff. from sex",
                          "Drink: eff. from smoke.",
                          "Smoking rate 1",
                          "Smoke. linear",
                          "Smoke. quadratic",
                          "Avg. Alter Smoke.",
                          "Smoke: eff. from sex",
                          "Smoke: eff. from drink.")[27:1])



g1 <- c("Friend rate 1",
                          "Density",
                          "Reciprocity",
                          "transTrip",
                          "transRecTrip",
                          "3-cycles",
                          "Indeg. Pop. sqrt.",
                          "Indeg. Act. sqrt.",
                          "Same sex")


g2 <- c("Alter drinking",
                          "Ego drinking",
                          "Ego x alt drinking",
                          "Alter smokimg",
                          "Ego smoking",
                          "Ego x alt smoking")
g3 <- c("Drinking rate 1",
                          "Drink. linear",
                          "Drink. quadratic",
                          "Avg. Alter Drink.",
                          "Drink: eff. from sex",
                          "Drink: eff. from smoke.",
                          "Smoking rate 1",
                          "Smoke. linear",
                          "Smoke. quadratic",
                          "Avg. Alter Smoke.",
                          "Smoke: eff. from sex",
                          "Smoke: eff. from drink.")

ld$g[ld$parameter %in% g1] <- "Structure"
ld$g[ld$parameter %in% g2] <- "Selection"
ld$g[ld$parameter %in% g3] <- "Behavior"

ld$g <- factor(ld$g, levels = c("Structure","Selection","Behavior"))

library(ggplot2)
p5 <- ggplot(ld, aes(x = parameter, y = Theta, group = model, colour = model,
                     shape = model)) +
         geom_point( size = 2.2 , position = position_dodge(width = .75)) +
         geom_hline(yintercept = 0, size = 1, color = "black") +
         geom_errorbar(size = .7, aes(ymin = Theta - SE, ymax = Theta + SE),
                        position = position_dodge(width = .75)) +
         theme_classic() +
         theme(axis.text.x = element_text(angle = 0, hjust = 1),
                axis.text.y = element_text(angle = 0, hjust = 1),
                legend.title = element_blank(),
                legend.position = "bottom",
                legend.direction = "vertical",
                legend.key.height = unit(1,"line"),
                legend.key.width = unit(1,"line"),
                text = element_text(size = 20),
                axis.title.x = element_blank()) +
         ylab("") + xlab("") + coord_flip()
p5


p6 <- ggplot(ld, aes(x = parameter, y = Theta, group = model, colour = model,
                     shape = model)) + scale_color_manual(values = c(
                    "grey75","grey60","grey45","grey30","grey15","grey0" )) +
  facet_wrap(facets = ~g, scales = "free", nrow = 3) +
  geom_point( size = 2.2 , position = position_dodge(width = .75)) +
  geom_hline(yintercept = 0, size = .25, color = "black") +
  geom_errorbar(size = .7, aes(ymin = Theta - SE, ymax = Theta + SE),
                position = position_dodge(width = .75)) +
  theme_classic() + theme(
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.text.y = element_text(angle = 0, hjust = 1),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(3,"line"), legend.key.width = unit(3,"line"),
    text = element_text(size = 20),
    axis.title.x = element_blank()) + ylab("") + xlab("") + coord_flip()
p6


## 2 standard errors
p6 <- ggplot(ld, aes(x = parameter, y = Theta, group = model, colour = model,
                     shape = model)) +
  facet_wrap(facets = ~g, scales = "free", nrow = 1) +
  geom_point( size = 2.2 , position = position_dodge(width = .75)) +
  geom_hline(yintercept = 0, size = .25, color = "black") +
  geom_errorbar(size = .7, aes(ymin = Theta - SE*2, ymax = Theta + SE*2),
                position = position_dodge(width = .75)) +
  theme_classic() + theme(
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.text.y = element_text(angle = 0, hjust = 1),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.height = unit(3,"line"), legend.key.width = unit(3,"line"),
    text = element_text(size = 20),
    axis.title.x = element_blank()) + ylab("") + xlab("") + coord_flip()
p6

```